The Effect of Weather Shocks on Children's Education in Bolivia

### 2020-05-10 final_seminar_paper.Rmd
at the point where everything's good except tables are somehow now after the discussion section instead of before and the discussion section is not showing up as a section, # visible

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F, 
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

@@

```{r import household data 2005, echo = F, message = F}
data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
  rename(hh_id = folio,
        assist1 = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s3_01,
        register1 = s3_04,
        school_level = s3_05a,
        grade = s3_05b,
        repeated = s3_07,
        income = y_perc1A,
        educ = aoesc,
        job = s4_16cod,
        tasks = s4_16b,
        pov_line = z,
        extr_pov = z_ext)

@@

```{r import household data 2006, echo = F, message = F}
data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
  rename(hh_id = Folio,
        assist1 = s4_08,  
        reason = S4_09, 
        members = mhogar,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s4_01,
        register1 = s4_04,
        school_level = s4_05a,
        grade = s4_05b,
        repeated = s4_07,
        income = yhogpcf,
        urb_rur = Urb_Rur,
        depto = ID01,
        educ = añoe,
        job = s5_17cod,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r import household data 2007, echo = F, message = F}
data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_10,  
        reason = s4_11, 
        members = mhogar,
        sex = s1_03,
        age = s1_04,
        relation = s1_06,
        literacy1 = s4_01,
        register1 = s4_06,
        school_level = s4_07a,
        grade = s4_07b,
        repeated = s4_09,
        income = yhogpcf,
        voucher = s4_05,
        educ = a_oe,
        job = s5_16co,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r import household data 2011, echo = F, message = F}
data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_05,
        grade = s4_06b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_06a,
        urb_rur = area,
        breakfast = s4_07,
        educ = e,
        job = cod_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r household data 2012, echo = F, message = F}
data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_04,
        grade = s4_05b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_05a,
        urb_rur = area,
        depto = departamento,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cods5_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r combine household datasets, echo = F, message = F}
all_years <- hh_2005 %>% 
  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)

@@

```{r import flood data, echo = F, message = F}
flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
  rename(dept = NOM_DEP)

@@

```{r importing droughts data, echo = F, message = F}
#drought data set 
library(readxl)
drought <- read_excel(file.path("Raw data", "recurrenciasequia_2002_2012bol (1).xls")) %>% 
#rename variables
    rename(dept = NOM_DEP)

@@

``` {r merging datasets, echo = F, message = F}
total_data<- all_years %>% 
  left_join(flood_total) %>%
  left_join(drought_total) %>%
  mutate(child = case_when(relation == 3 ~ 1,
                           relation != 3 ~ 0),
         grandchild = case_when(relation == 8 ~ 1,
                                relation != 8 ~ 0),
         pov_diff = income - pov_line,
         pov = case_when(pov_diff > 0 ~ 0,
                         pov_diff < 0 ~ 1),
         extr_diff = income - extr_pov,
         extr = case_when (extr_diff > 0 ~ 0,
                           extr_diff < 0 ~ 1),
                  work_reason = case_when(reason == "work" ~ 1,
                                 reason != "work" ~ 0),
         money_reason = case_when(reason == "money" ~ 1, 
                                  reason != "money" ~ 0))

@@

```{r creating head of household subset, echo = F, message = F}

@@

```{r creating children subset, echo = F, message = F}

@@

total_children %>%
      group_by(child, grandchild) %>%
      summarise(n = n())

total_children %>%
   group_by(work_reason, money_reason) %>%
   summarise(n = n())

@@

```{r combining household head characteristics with children dataset, echo = F, message = F}
hh_head_2 <- hh_head %>%
  select(female, age, literacy, educ, job, tasks, income, pov, extr, hh_id) %>%
  rename (female_hhhead = female,
          hhhead_age = age,
          hhhead_literacy = literacy,
          hhhead_educ = educ,
          hhhead_job = job,
          hhhead_tasks = tasks)

@@

total %>%
  group_by(reason) %>%
  summarise(n = n())

@@

```{r regression variable specifications, echo = F, message = F}

@@

reg_data %>%
  group_by(dept_factor) %>%
  summarise(n = n())

reg_data %>%
   group_by(attend, register) %>%
   summarise(n = n())

```
##Variable Descriptions

@@

```{r create summary statistic table, echo = F, results = 'asis'}
#creating a table with the main explanatory variables

@@

```{r probit regression total sample, echo = F, message = F}

@@

```{r generate probit regression table, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
               title = "Full sample probit regression results", 
               label = "fullResults1",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression rural sample, echo = F, message = F}

@@

```{r generate probit regression table rural sample, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_r, probitMod_register_r, probitMod_work_r), 
               title = "Rural sample probit regression results", 
               label = "fullResults3",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression urban sample, echo = F, message = F}

@@

The following tables show the results with the lagged variables for floods and droughts.
```{r probit regression total sample lagged variables, echo = F, message = F}

@@

```{r generate probit regression table lagged variables, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lt, probitMod_register_lt, probitMod_work_lt), 
               title = "Full sample probit regression results (lagged)", 
               label = "fullResults5",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression rural sample lagged, echo = F, message = F}

@@

```{r generate probit regression table rural sample lagged, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lr, probitMod_register_lr, probitMod_work_lr), 
               title = "Rural sample probit regression results (lagged)", 
               label = "fullResults6",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

##Robustness Checks

@@

```{r probit regression total sample extr, echo = F, message = F}

@@

```{r generate probit regression table extr, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
               title = "Full sample probit regression results (poverty measure)", 
               label = "fullResults8",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

The following results show the linear regression models.

@@

```{r generate regression table, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
               title = "Full sample linear regression results", 
               label = "fullResults11",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
```

@@

>>>>

```{r setup, echo = F, warning = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F, 
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

@@

```{r import household data 2005, echo = F, warning = F, message = F}
data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
  rename(hh_id = folio,
        assist1 = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s3_01,
        register1 = s3_04,
        school_level = s3_05a,
        grade = s3_05b,
        repeated = s3_07,
        income = y_perc1A,
        educ = aoesc,
        job = s4_16cod,
        tasks = s4_16b,
        pov_line = z,
        extr_pov = z_ext)

@@

```{r import household data 2006, echo = F, warning = F, message = F}
data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
  rename(hh_id = Folio,
        assist1 = s4_08,  
        reason = S4_09, 
        members = mhogar,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s4_01,
        register1 = s4_04,
        school_level = s4_05a,
        grade = s4_05b,
        repeated = s4_07,
        income = yhogpcf,
        urb_rur = Urb_Rur,
        depto = ID01,
        educ = añoe,
        job = s5_17cod,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r import household data 2007, echo = F, warning = F, message = F}
data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_10,  
        reason = s4_11, 
        members = mhogar,
        sex = s1_03,
        age = s1_04,
        relation = s1_06,
        literacy1 = s4_01,
        register1 = s4_06,
        school_level = s4_07a,
        grade = s4_07b,
        repeated = s4_09,
        income = yhogpcf,
        voucher = s4_05,
        educ = a_oe,
        job = s5_16co,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r import household data 2011, echo = F, warning = F, message = F}
data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_05,
        grade = s4_06b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_06a,
        urb_rur = area,
        breakfast = s4_07,
        educ = e,
        job = cod_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r household data 2012, echo = F, warning = F, message = F}
data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_04,
        grade = s4_05b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_05a,
        urb_rur = area,
        depto = departamento,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cods5_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r combine household datasets, echo = F, warning = F, message = F}
all_years <- hh_2005 %>% 
  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)

@@

```{r import flood data, echo = F, warning = F, message = F}
flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
  rename(dept = NOM_DEP)

@@

```{r importing droughts data, echo = F, warning = F, message = F}
#drought data set 
library(readxl)
drought <- read_excel(file.path("Raw data", "recurrenciasequia_2002_2012bol (1).xls")) %>% 
#rename variables
    rename(dept = NOM_DEP)

@@

``` {r merging datasets, echo = F, warning = F, message = F}
total_data<- all_years %>% 
  left_join(flood_total) %>%
  left_join(drought_total) %>%
  mutate(child = case_when(relation == 3 ~ 1,
                           relation != 3 ~ 0),
         grandchild = case_when(relation == 8 ~ 1,
                                relation != 8 ~ 0),
         pov_diff = income - pov_line,
         pov = case_when(pov_diff > 0 ~ 0,
                         pov_diff < 0 ~ 1),
         extr_diff = income - extr_pov,
         extr = case_when (extr_diff > 0 ~ 0,
                           extr_diff < 0 ~ 1),
                  work_reason = case_when(reason == "work" ~ 1,
                                 reason != "work" ~ 0),
         money_reason = case_when(reason == "money" ~ 1, 
                                  reason != "money" ~ 0))

@@

```{r creating head of household subset, echo = F, warning = F, message = F}

@@

```{r creating children subset, echo = F, warning = F, message = F}

@@

@@

```{r combining household head characteristics with children, echo = F, warning = F, results = 'hide', message = F}
hh_head_2 <- hh_head %>%
  select(female, age, literacy, educ, job, tasks, income, pov, extr, hh_id) %>%
  rename (female_hhhead = female,
          hhhead_age = age,
          hhhead_literacy = literacy,
          hhhead_educ = educ,
          hhhead_job = job,
          hhhead_tasks = tasks)

@@

total %>%
  group_by(reason) %>%
  summarise(n = n())
#figure out how to get rid of the table

@@

```{r regression variable specifications, echo = F, warning = F, message = F}

@@

##Variable Descriptions

@@

```{r create summary statistic table, echo = F, message = F, warning = F, results = 'asis'}
#creating a table with the main explanatory variables

@@

```{r probit regression total sample, echo = F, warning = F, message = F}

@@

```{r generate probit regression table, echo = F, message = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
               title = "Full sample probit regression results", 
               label = "fullResults1",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression rural sample, echo = F, warning = F, message = F}

@@

```{r generate probit regression table rural sample, echo = F, message = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_r, probitMod_register_r, probitMod_work_r), 
               title = "Rural sample probit regression results", 
               label = "fullResults3",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression urban sample, echo = F, warning = F, message = F}

@@

The following tables show the results with the lagged variables for floods and droughts.

```{r probit regression total sample lagged variables, echo = F, warning = F, message = F}

@@

```{r generate probit regression table lagged variables, echo = F, message = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lt, probitMod_register_lt, probitMod_work_lt), 
               title = "Full sample probit regression results (lagged)", 
               label = "fullResults5",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

```{r probit regression rural sample lagged, echo = F, warning = F, message = F}

@@

```{r probit regression table rural sample lagged, message= F, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lr, probitMod_register_lr, probitMod_work_lr), 
               title = "Rural sample probit regression results (lagged)", 
               label = "fullResults6",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

## Robustness Checks

@@

```{r probit regression total sample extr, echo = F, warning = F, message = F}

@@

```{r generate probit regression table extr, echo = F, message = F, results = "asis"}
stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
               title = "Full sample probit regression results (poverty measure)", 
               label = "fullResults8",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

@@

@@

```{r generate regression table, echo = F, message = F, results = "asis"}
stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
               title = "Full sample linear regression results", 
               label = "fullResults11",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
```

@@

# References
Baez, J.E., Lucchetti, I., Genoni, M. E., & Salazar, M. (2017). Gone with the storm: Rainfall shocks and household wellbeing in Guatemala. Journal of Development Studies, 53(8), 1253-1271.

Groppo, V., and K. Kraehnert. 2017. The Impact of Extreme Weather Events on Education. Journal of Population Economics 30 (2): 433–72

Jensen, R. (2000). Agricultural volatility and investments in children. American Economic Review Paper and Proceedings, 90(2), 399-404.

Marchetta, F., Sahn, D., Tiberti, L. (2019). The Role of Weather on Schooling and Work of Young Adults in Madagascar. American Journal of Agricultural Economics, 101(4), 1203-1227.

Shah, M., & Steinberg, B.M. (2017). Drought of opportunities: Contemporaneous and long-term impacts of rainfall shocks on human capital. Journal of Political Economy, 125(2), 527-561.

Zimmermann, L. (2020). Remember When It Rained—Schooling Responses to Shocks in India. World Development, 126.
